Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex
various edits
removing DNF grammar
savin' space

\usepackage{tikz}
\usepackage{xypic}
%\usepackage{natbib}

@@

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
then $w$ is said to \emph{generalise} $v$
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$ defined:
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
where $\preceq_n$ is a reflexive, transitive relation on
$(\mathbb{Z}_{\infty})^n \times (\mathbb{Z}_{\infty})^n$. Thus, the
element $\infty$ generalises any index in the same dimension.
%
\end{definition}

\begin{example}
Consider the following array subscript to indexing schemes
conversions:
%
\begin{align*}
\begin{array}{ll}
\schemeN{(i, j+1)} & =\vtwoh{0}{1} \\
\schemeN{(1, j+1)} & = \vtwoh{\infty}{1} \\
\schemeN{(3, j)} & = \vtwoh{\infty}{0} \\
\schemeN{(i*2, j)} & = \bot
\end{array}
\end{align*}
There is one generalisation between these
schemes, namely that $\vtwoh{0}{1} \preceq_2 \vtwoh{\infty}{1}$.
The last example is undefined as its first expression is a
non-neighbour relative index.
\end{example}

@@

The notion of which dimensions $i$ have \emph{any scheme in the model
where $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
\begin{definition}%[Constrained dimensions]
For an $n$-dimensional model $M$, its \emph{constrained dimensions}
are the set of dimension identifiers where at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}
\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is:
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.2cm}{$(u, v) \in M \times N,$ \\[0.1em]
                  $\;\; i \in \{1, \ldots, 2^n\}, \,
                  x = (u \bowtie v)_i,$ \\[0.1em]
                  $\;\; j \in (\effdims{M}{n} \cup
                   \effdims{N}{n}),$\\[0.1em]
                   $\;\; x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The first guard takes the Cartesian product of
the models and quantifies over all pairings $(u, v)$.
The second guard binds $j$ which ranges over the constrained
dimensions for $M$ unioned with the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constrained.
The \emph{pairwise permutation} $u \bowtie v$
builds a $2^n \times n$ matrix
of all possible $n$-vectors generated by
non-deterministically picking for each $j^{th}$ entry either
$u_j$ or $v_j$ \eg{}
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
correspond to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. The pairwise permutation
matrix is thus defined in terms of the logical
matrix $b$ where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$ as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
The permuted schemes $x = (u \bowtie v)_i$ added to the model
$M \otimes N$ are only those without $\infty$ in
the constrained dimensions given by each model ($x_j \neq
\infty$). This eliminates all permutations with
$\infty$ in a constrained dimension position.
\end{definition}

@@

\item (\emph{overlapping}) Given two region constants $R_1$ and $R_2$
on the same dimension, then some combinations of region constants
can be coalesced into a single region constant $R_3$ if $R_1$ completely
overlaps $R_2$. There are five possible instantiations of the
following rule $R_1 + R_2 \equiv R_3$:
%
\newcommand{\stenGR}[4]{\small{\textit{#4}(\term{depth=}#1,
  \term{dim=}#2, #3)}}
\newcommand{\stenG}[3]{\small{\textit{#3}(\term{depth=}#1,
  \term{dim=}#2)}}
\begin{align*}
\hspace{-1em} & \stenGR{n}{d}{p \vee q}{$R_3$} \\[-0.4em]
\hspace{-1em} \equiv \; & \stenGR{n}{d}{p}{$R_1$} \; \texttt{+} \; \stenGR{m}{d}{q}{$R_2$}
\end{align*}
where $n \geq m$ and $p$ and $q$ range
over booleans representing the absence (true) or presence (false)
of \irrefl{} which are combined via disjunction.
We denote the possibilities as $\term{fwd+fwd} \equiv \term{fwd}$, $\term{bwd+bwd} \equiv \term{bwd}$,
$\term{cen+cen} \equiv \term{cen}$, $\term{cen+fwd} \equiv \term{cen}$ and
$\term{cen+bwd} \equiv \term{cen}$. %Commutative of $\term{+}$
%provides symmetric variants.

\item (\textit{overlapping pointed}) \term{pointed} regions coalesce with other regions of the
same dimension:
%
\begin{align*}
& \stenG{n}{d}{region} \\[-0.4em]
\equiv \; & \stenGR{n}{d}{p}{region} \; \texttt{+} \;
                          \stenReflS{d}
\end{align*}
%
\item \textit{(centered)} a forward and backward region of the same
  depth and dimension coalseces into a centered region:
\begin{align*}
\hspace{-1em} & {\small{\stenCen{n}{d}{, p \vee q}}} \\[-0.4em]
\hspace{-1em} \equiv \; & {\small{\stenFwdSR{n}{d}{p} \; \texttt{+} \;
                          \stenBwdSR{n}{d}{q}}}
\end{align*}
%%
\item $(\textsc{point}) \;\; (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \; \equiv \; R
                         \, \term{*} \, S$
meaning that for the sum of products of two regions $R$ and $S$,
where in each component one of $R$ or $S$ has the non-pointed
attribute denoted by $R^{\term{np}}$ and $S^{\term{np}}$,
the non-pointed attributes cancel to give $R \, \term{*} \, S$.
%
\item $(\textsc{$\ast$Dim}) \;\; R^d \, \term{$\ast$} \, S^d \equiv R^d
                             \, \term{+} \, S^d$
%
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
Thus a product of two regions on
the same dimension is equal to their addition.
\end{itemize}
%
Stencil specifications in \textit{spec} are considered
equal when they have the same modifiers and $\equiv$ equates
their regions.

@@

\noindent
We outline here the procedures for checking conformance
of source code against specifications (\Cref{subsec:checking})
and for infering specifications from code (\Cref{subsec:inference}).
Both rely on a program analysis that converts array indexing
expressions into sets of index schemes. We outline this analysis
first (\cref{subsec:analysis}). Note that the analysis procedure
can be made arbitrarily more complex and wide-ranging independent
of the checking and inference procedures. At the moment, the analysis
is largely \emph{syntactic}, with only a small amount of
semantic interpretation of the code.

@@

\begin{figure}[t]
\vspace{-1em}
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
\trule{up}\dfrac{\forall u \in M, \exists v \in N .  \, u \preceq v}
      {\cons{M}{\textsf{upper}(N)}}
\,
\trule{low}\dfrac{\forall v \in N, \exists u \in M . \, u \preceq v}
      {\cons{M}{\textsf{lower}(N)}} \\[1.2em]
%%
\trule{exact}\dfrac{\cons{M}{\textsf{lower}(N)} \qquad \cons{M}{\textsf{upper}(N)}}
  {\cons{M}{\textsf{exact}(N)}} \\[1.2em]
%%
\trule{both}\dfrac{\cons{M}{\textsf{lower}(M_1)} \qquad \cons{M}{\textsf{upper}(M_2)}}
  {\cons{M}{\textsf{both}(M_1, M_2)}}\\[1.2em]
%%
\trule{$1$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{once}(N)}}
\;\;
\trule{$*$}
\dfrac{\cons{M}{N}}{\consA{\textsf{mult}(M)}{\textsf{mult}(N)}}
  \\[1.2em]
\trule{$1 < *$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{mult}(N)}}
\end{array}
\end{align*}
\vspace{-0.5em}
\caption{Consistency of models (code vs. specification)}
\label{fig:consistency}
\vspace{-0.9em}
\end{figure}

@@

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see \Cref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec},
containing specifications in disjunctive normal form over regions
where the disjunct is \term{+} and the conjunct is \term{*}; 
 elements of \specDNF{} are of the form:
%%
$
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
$.

More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%
\begin{align*}
\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \quad\;\;
\tySum{\specDNF} ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
In the following, $S, T$ range over elements of
\specDNF{} and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$, \ie{} $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$.
\end{definition}

@@

\begin{definition} The $\odot$ operation on $\textsf{Approx}(\specDNF{})$
  is defined in terms of the intermediate operator $\hat{\odot}$ on
  \specDNF{}, defined:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form, where $\hat{\odot}$ is essentially applying
 the \trule{\textsc{dist}} rule of the equational theory
 (\Cref{fig:equations}). The operation is
commutative and associative, and sound with respect to our
model (see \Cref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
       \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u)  = \; & \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u) \\
\textit{inj}(S) \odot \textit{inj}(T) = \; & \textit{inj}(S
                                              \hat{\odot} T)
\end{align*}
where in the last case $\textit{inj}$ corresponds to the unary
injections of \textsf{Approx}. For brevity we omit the cases where the arguments above are flipped
as $\odot$ is commutative.
\end{definition}

@@

\section{Evaluation}

>>>>

\usepackage{tikz}
\usepackage{xypic}
\usepackage{natbib}

@@

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
then $w$ is said to \emph{generalise} $v$
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$ defined:
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
where $\preceq_n$ is a reflexive, transitive relation on
index schemes $\domainVal^n$. Thus,
 $\infty$ generalises any index in the same dimension.
%
\end{definition}

\begin{example}
Consider the following array subscript to indexing schemes
conversions:
%
\begin{align*}
\begin{array}{ll}
\schemeN{(i, j+1)} & =\vtwoh{0}{1} \\
\schemeN{(1, j+1)} & = \vtwoh{\infty}{1} \\
\schemeN{(i*2, j)} & = \bot
\end{array}
\end{align*}
There is one generalisation between these
schemes, namely that $\vtwoh{0}{1} \preceq_2 \vtwoh{\infty}{1}$.
The last example is undefined as its first expression is a
non-neighbour relative index.
\end{example}

@@

The notion of which dimensions $i$ have \emph{any scheme in the model
where $u_i \neq \infty$} is represented by following  function:
\begin{definition}%[Constrained dimensions]
For an $n$-dimensional model $M$, its \emph{constrained dimensions}
are the set of dimension identifiers where at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}
\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is:
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.2cm}{$(u, v) \in M \times N,$ \\[0.1em]
                  $\;\; i \in \{1, \ldots, 2^n\}, \,
                  x = (u \bowtie v)_i,$ \\[0.1em]
                  $\;\; j \in (\effdims{M}{n} \cup
                   \effdims{N}{n}),$\\[0.1em]
                   $\;\; x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The first guard takes the Cartesian product of
the models and quantifies over all pairings $(u, v)$.
The second guard binds $j$ which ranges over the constrained
dimensions for $M$ unioned with the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constrained.
The \emph{pairwise permutation} $u \bowtie v$
builds a $2^n \times n$ matrix
of all possible $n$-vectors generated by
non-deterministically picking for each $j^{th}$ entry either
$u_j$ or $v_j$ \eg{}
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
correspond to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. The pairwise permutation
matrix is thus defined in terms of the logical
matrix $b$ where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$ as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
The permuted schemes $x = (u \bowtie v)_i$ added to the model
$M \otimes N$ are only those without $\infty$ in
the constrained dimensions given by each model ($x_j \neq
\infty$). This eliminates all permutations with
$\infty$ in a constrained dimension position.
\end{definition}

@@

\item (\emph{overlapping}) Given two region constants $R_1$ and $R_2$
on the same dimension, then some combinations of region constants
can be coalesced into a single region constant $R_3$ if $R_1$ completely
overlaps $R_2$. There are five possible instantiations of the
following rule $R_1 + R_2 \equiv R_3$:
%
\newcommand{\stenGR}[4]{\small{\textit{#4}(\term{depth=}#1,
  \term{dim=}#2, #3)}}
\newcommand{\stenG}[3]{\small{\textit{#3}(\term{depth=}#1,
  \term{dim=}#2)}}
%
\vspace{-0.2em}
\begin{align*}
\hspace{-1em} & \stenGR{n}{d}{p \vee q}{$R_3$} \\[-0.4em]
\hspace{-1em} \equiv \; & \stenGR{n}{d}{p}{$R_1$} \; \texttt{+} \; \stenGR{m}{d}{q}{$R_2$}
\end{align*}
where $n \geq m$ and $p$ and $q$ range
over booleans representing the absence (true) or presence (false)
of \irrefl{} which are combined via disjunction.
We denote the possibilities as $\term{fwd+fwd} \equiv \term{fwd}$, $\term{bwd+bwd} \equiv \term{bwd}$,
$\term{cen+cen} \equiv \term{cen}$, $\term{cen+fwd} \equiv \term{cen}$ and
$\term{cen+bwd} \equiv \term{cen}$. %Commutative of $\term{+}$
%provides symmetric variants.

\item (\textit{overlapping pointed}) \term{pointed} regions coalesce
  with other regions of the same dimension:
%
\vspace{-0.4em}
\begin{align*}
& \stenG{n}{d}{region} \\[-0.4em]
\equiv \; & \stenGR{n}{d}{p}{region} \; \texttt{+} \;
                          \stenReflS{d}
\end{align*}
\vspace{-1.5em}

\item \textit{(centered)} a forward and backward region of the same
  depth and dimension coalseces into a centered region:
%
\vspace{-0.4em}
\begin{align*}
\hspace{-1em} & {\small{\stenCen{n}{d}{, p \vee q}}} \\[-0.4em]
\hspace{-1em} \equiv \; & {\small{\stenFwdSR{n}{d}{p} \; \texttt{+} \;
                          \stenBwdSR{n}{d}{q}}}
\end{align*}
\vspace{-1.5em}

\item $(\textsc{point}) \;\; (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \; \equiv \; R
                         \, \term{*} \, S$
meaning that for the sum of products of two regions $R$ and $S$,
where in each component one of $R$ or $S$ has the non-pointed
attribute denoted by $R^{\term{np}}$ and $S^{\term{np}}$,
the non-pointed attributes cancel to give $R \, \term{*} \, S$.
%
\item $(\textsc{$\ast$Dim}) \;\; R^d \, \term{$\ast$} \, S^d \equiv R^d
                             \, \term{+} \, S^d$
%
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
Thus a product of two regions on
the same dimension is equal to their addition.
\end{itemize}
%
Stencil specifications in \textit{spec} are considered
equal when they have the same modifiers and $\equiv$ equates
their regions.

@@

\noindent
We outline here the procedures for checking conformance
of source code against specifications (\Cref{subsec:checking})
and for infering specifications from code (\Cref{subsec:inference}).
Both rely on a program analysis that converts array subscripts
 into sets of index schemes. We outline this analysis
first (\cref{subsec:analysis}). Note that the analysis 
can be made arbitrarily more complex and wide-ranging independent
of the checking and inference procedures. At the moment, the analysis
is largely \emph{syntactic}, with only a small amount of
semantic interpretation of the code.

@@

\begin{figure}[t]
\vspace{-0.5em}
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
\trule{up}\dfrac{\forall u \in M, \exists v \in N .  \, u \preceq v}
      {\cons{M}{\textsf{upper}(N)}}
\,
\trule{low}\dfrac{\forall v \in N, \exists u \in M . \, u \preceq v}
      {\cons{M}{\textsf{lower}(N)}} \\[1.2em]
%%
\trule{exact}\dfrac{\cons{M}{\textsf{lower}(N)} \qquad \cons{M}{\textsf{upper}(N)}}
  {\cons{M}{\textsf{exact}(N)}} \\[1.2em]
%%
\trule{both}\dfrac{\cons{M}{\textsf{lower}(M_1)} \qquad \cons{M}{\textsf{upper}(M_2)}}
  {\cons{M}{\textsf{both}(M_1, M_2)}}\\[1.2em]
%%
\trule{$1$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{once}(N)}}
\;\;
\trule{$*$}
\dfrac{\cons{M}{N}}{\consA{\textsf{mult}(M)}{\textsf{mult}(N)}}
  \\[1.2em]
\trule{$1 < *$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{mult}(N)}}
\end{array}
\end{align*}
\vspace{-0.5em}
\caption{Consistency of models (code vs. specification)}
\label{fig:consistency}
\vspace{-0.9em}
\end{figure}

@@

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see \Cref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec},
containing specifications in disjunctive normal form over regions
where the disjunct is \term{+} and the conjunct is \term{*}; 
 elements of \specDNF{} are of the form:
%%
$
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
$.
%
%More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%%
%\begin{align*}
%\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \quad\;\;
%\tySum{\specDNF} ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
%  \tyProd{\specDNF} \\
%\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
%   \textit{region}
%\end{align*}
%%
In the following, $S, T$ range over elements of
\specDNF{} and $\tyProd{s}, \tyProd{t}$ range over conjuncts
 \eg{} $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$.
\end{definition}

@@

\begin{definition} The $\odot$ operation on $\textsf{Approx}(\specDNF{})$
  is defined in terms of the intermediate operator $\hat{\odot}$ on
  \specDNF{}, defined:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form, where $\hat{\odot}$ is essentially applying
 the \trule{\textsc{dist}} rule of the equational theory. The operation is
commutative and associative, and sound with respect to our
model (see \Cref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
       \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u)  = \; & \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u) \\
\textit{inj}(S) \odot \textit{inj}(T) = \; & \textit{inj}(S
                                              \hat{\odot} T)
\end{align*}
where in the last case $\textit{inj}$ corresponds to the unary
injections of \textsf{Approx}. For brevity we omit the cases where the arguments above are flipped
as $\odot$ is commutative.
\end{definition}

@@

\section{Evaluation}
\label{sec:evaluation}
